---
title: "GAE7"
author: "Hannah Alexander"
format: 
  html:
    theme: custom.scss
    embed-resources: true
    code-tools: true

csl: ecology.csl

editor: visual
bibliography: references.bib
---

```{r}
#| label: SetUp
#| message: FALSE
#| echo: FALSE

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
  ) 

library(emmeans)
library(lmerTest)
library(effectsize)
library(car)
library(ggfortify)
library(patchwork)
library(broom)
library(latex2exp)       
library(kableExtra)
library(tidyverse)

options(knitr.kable.NA = '')
```

## Introduction

@Pintar.etal2018 conducted a field mesocosm experiment to explore the relative effects of predation risk and nutrient availability (a pdf of their publication is included in the GitHub repository). The predation risk manipulation was accomplished by adding either 0, 1, or 2, Golden Topminnows (*Fundulus chrysotus*) to each mesocosm. These small fish are surface-feeding predators and are known to elicit colonization and oviposition avoidance among dytiscid beetles and *Culex* mosquitoes, but not in some hydrophilid beetles. Nutrient-level treatments were created by the addition of 0, 4, or 8 g of rabbit chow (Small World Rabbit Food, Mannapro, St. Louis, MO, 40% protein).

## Methods

Please see @Pintar.etal2018 for greater detail on experimental design and data collection. In brief, the experiment was set up as a 3 X 3 factorial treatment design (9 treatment combinations). The experiment was replicated in 3 separate locations at 3 different times for a total of 9 replicates. This should have resulted in 81 observations total, however some fish died during the experiment rendering those observations not being included in the data resulting in only 69 observations.

```{r }
#| label: dataImport

df1 <- read_csv("Pintar_etal_2018.csv",
                show_col_types = FALSE)

# Block = 1 to 9 for the 9 spatiotemporal blocks (3 locations X 3 times)
# Pool = 1 to 9 for the plastic wading pools used for the experiment
# Location = 1 to 3 for the 3 fields where the pools were deployed
# Time = a to c for the 3 time periods used for each 2-week run of the exp.
# Nutrient = 1 to 3 for the 3 nutrient levels
# Fish = 1 to 3 for the 3 fish densities
# Culex_Egg_Rafts = number of egg rafts collected over the 2 weeks
# Hydrophilidae = number of hydrophilid beetles (Water Scavenger Beetles)
# Dytiscidae = number of dytiscid beetles (Predaceous Diving Beetles)

# Fish and Nutrient are provided as integers, but I wish to use them as ordered
# factors so that, by default, polynomial contrasts will be used to compute the
# ANOVAs.

df1$Nutrient <- factor(df1$Nutrient,
                       levels = c("0", "4", "8"),
                       ordered = TRUE)

df1$Fish <- factor(df1$Fish,
                   levels = c("0", "1", "2"),
                   ordered = TRUE)

df1$Block <- factor(df1$Block)

```

### Statistical Methods

I ran a linear mixed-effect model for the three organisms of interest, Culex mosquitoes, Hydrophilid beetles, and Dytiscid beetles using the lmer function from the [@lme4] package. Blocks were treated as a random effect. All tables were produced using the [@tidyverse] package.

## Results

Availability of nutrients, amount of fish, and block had a significant effect on the number of Culex mosquito eggs @tbl-mixedANOVA-Culex, but the interaction between nutrients and fish did not have a statistically significant effect. Similar results were seen with Dytiscid beetles, with the only parameter not having a statistically significant effect being the interaction between nutrients and fish @tbl-mixedANOVA-Dyt. Conversely, the only parameter to have a significant effect on the abundance of Hydrophilid beetles was the effect of the block @tbl-mixedANOVA-Hydro.

```{r}
#| label: tbl-mixedANOVA-Culex
#| tbl-cap: "Linear mixed-effect model results for effect of Nutrients, Fish, and spatiotemporal blocks on Culex eggs."
#| tbl-subcap: 
#|     - "Fixed effect"
#|     - "Random effect"
#| layout-nrows: 2

CulexmodMixed <- lmer(Culex_Egg_Rafts ~ Nutrient*Fish + (1|Block), data = df1)

CulexoutMixed <- anova(CulexmodMixed)

kbl(
  CulexoutMixed,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "P"
    ),
  digits = c(3,3,0,0,3,4)
  ) |>
  kable_classic()

kbl(
  rand(CulexmodMixed),
  col.names = c(
    "Parameters",
    "log-likelihood",
    "AIC",
    "LRT",
    "df",
    "P"
    ),
  digits = c(0,3,3,3,0,4)
  )
```

```{r}
#| label: tbl-mixedANOVA-Hydro
#| tbl-cap: "Linear mixed-effect model results for effect of Nutrients, Fish, and spatiotemporal blocks on Hydrophilid beetles."
#| tbl-subcap: 
#|     - "Fixed effect"
#|     - "Random effect"
#| layout-nrows: 2

HydromodMixed <- lmer(Hydrophilidae ~ Nutrient*Fish + (1|Block), data = df1)

HydrooutMixed <- anova(HydromodMixed)

kbl(
  HydrooutMixed,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "P"
    ),
  digits = c(3,3,0,0,3,4)
  ) |>
  kable_classic()

kbl(
  rand(HydromodMixed),
  col.names = c(
    "Parameters",
    "log-likelihood",
    "AIC",
    "LRT",
    "df",
    "P"
    ),
  digits = c(0,3,3,3,0,4)
  )
```

```{r}
#| label: tbl-mixedANOVA-Dyt
#| tbl-cap: "Linear mixed-effect model results for effect of Nutrients, Fish, and spatiotemporal blocks on Dytiscid beetles."
#| tbl-subcap: 
#|     - "Fixed effect"
#|     - "Random effect"
#| layout-nrows: 2

DytmodMixed <- lmer(Dytiscidae ~ Nutrient*Fish + (1|Block), data = df1)

DytoutMixed <- anova(DytmodMixed)

kbl(
  DytoutMixed,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "P"
    ),
  digits = c(3,3,0,0,3,4)
  ) |>
  kable_classic()

kbl(
  rand(DytmodMixed),
  col.names = c(
    "Parameters",
    "log-likelihood",
    "AIC",
    "LRT",
    "df",
    "P"
    ),
  digits = c(0,3,3,3,0,4)
  )
```

## References
